\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}		% lets you input equations in math mode
\usepackage{graphicx}		% lets you include images
\usepackage{enumerate}		% lets you make lists
\usepackage{hyperref}		% lets you make links
\usepackage{subcaption}     % if you want to use subcaptions
\usepackage[all]{hypcap}	% makes links refer to figures and not captions
\usepackage{relsize}		% lets you use relative font sizes
\usepackage{caption}        % lets you add captions
\usepackage{array}          % lets you specify table column widths
\usepackage[margin=1in, paperwidth=8.5in, paperheight=11in]{geometry} % I'll bet you can figure this one out
\usepackage[utf8]{inputenc} % character encoding - you don't need to understand this
\usepackage{circuitikz}
\usepackage[utf8]{inputenc}

\title{Linearity One Final Project Proposal}
\author{Alison Palmer and Anusha Datar}
\date{April 25, 2018}

\begin{document}

\maketitle

\section{Proposed Project Outline}
Over the course of this final project, we hope to learn more about signal processing and the Fourier Transform based on the linear algebra and differential equations material covered over the course of Linearity One. Overall, our learning goals are to understand:

\begin{enumerate}
  \item The mathematical background and computation of the Discrete Fourier Transform (DFT) and its applications.
  \item The mathematical background and computation of the Fast Fourier Transform (FFT) Algorithm and its applications.
  \item How to use the Fast Fourier Transform with real signals and software, likely with guitar chord recognition. 
\end{enumerate}

\section{Proposed Material and Resources}

Some research papers, videos, and other publications we plan to use for this project include (note that these are links):

\begin{enumerate}
    \item\href {http://ijcsi.org/papers/IJCSI-11-3-1-139-144.pdf}{Chord Identification Chord Identification Using Pitch Class Profile Method with Using Pitch Class Profile Method with Fast Fourier Transform Feature Extraction}
    \item\href{http://www.fim.uni-passau.de/fileadmin/files/lehrstuhl/sauer/geyer/BA_MA_Arbeiten/BA-HausnerChristoph-201409.pdf}{Design and Evaluation of a Simple Chord
    Detection Algorithm} 
    \item\href{http://www.math.uchicago.edu/~may/VIGRE/VIGRE2010/REUPapers/Anand.pdf}{A Brief Study of Discrete and Fast Fourier Transforms}
    \item\href{https://www.youtube.com/watch?v=M0Sa8fLOajA}{Gilbert Strang's Fast Fourier Transform Lecture}
    

\end{enumerate}

\section{Deliverables}
Through this project, we hope to expand our current understanding of mathematics through expanding both our theoretical understanding and developing a project. For that reason, our proposed deliverables include:
\begin{itemize}
  \item A four to five page report covering what we learned and how we developed our project.
  \item A video lasting no longer than five minutes that includes explanation of the concepts we learned over the course of this project.
  \item A software repository and demonstration.
\end{itemize}


\end{document}
